Wired for Success: Sweden's Electrical Innovators in the Age of Industrial Transformation

### 2023-11-20 index.qmd
add examples

---
title: "Wired for Success: Sweden's Electrical Innovators in the Age of Industrial Transformation"
subtitle: "Labour Mobility Lab in Lund"
date: today
author: "Jonathan Jayes  - Lund University"
title-slide-attributes:
  data-background-image: "assets/preview.png"
  data-background-size: "30%"
  data-background-opacity: "0.2"
  data-background-position: "right"
  data-background-repeat: "repeat"
format:
  clean-revealjs:
    code-overflow: wrap
    width: 1600
    height: 900
    fig-align: center
    default-image-extension: svg
    footer: "[{{< fa house >}}](https://interludeone.com)"
    slide-number: c
    code-block-bg: true
    code-block-border-left: "#31BAE9"
image: "preview_img/03-coordinate-systems-and-axes.jfif"
favicon: html/favicon.ico
---

@@

skills_table %>%
    tab_options(table.font.size = 40)
```

@@

I spent months doing this last year to classify occupations to HISCLASS groups, hand labelling data and optimizing a support vector machine model.

@@

```{r}
knitr::include_graphics("assets/svm.png")
```

@@

### We make use of pre-trained large language models
- Benefit from semantic similarity
- Text embeddings mean we can have an arbitrary number of classes

@@

```{r}
knitr::include_graphics("assets/umap_1.png")
```

@@

@@

### 026 Metallurgists

@@

### Karl Lund in 1923

@@

@@

## Matricies of HISCO codes

```{r}
knitr::include_url("assets/heatmap.html", height = "800")

>>>>

---
title: "Wired for Success: Sweden's Electrical Innovators in the Age of Industrial Transformation"
subtitle: "Labour Mobility Lab in Lund"
date: today
author: "Jonathan Jayes  - Lund University"
title-slide-attributes:
  data-background-image: "assets/preview.png"
  data-background-size: "30%"
  data-background-opacity: "0.2"
  data-background-position: "right"
format:
  clean-revealjs:
    code-overflow: wrap
    width: 1600
    height: 900
    fig-align: center
    default-image-extension: svg
    footer: "[{{< fa house >}}](https://interludeone.com)"
    slide-number: c
    code-block-bg: true
    code-block-border-left: "#31BAE9"
image: "preview_img/03-coordinate-systems-and-axes.jfif"
favicon: html/favicon.ico
---

@@

skills_table %>%
    tab_options(table.font.size = 32)
```

@@

### Building one labelled classification scheme manually does not generalize easily
- If you code all of your occupations to HISCO and then want to move to SSYK, you have to redo all of the codings or try and find a crosswalk.

::: {.column width="40%"}
I spent months doing this last year to classify occupations to HISCLASS groups, hand labelling data and optimizing a support vector machine model.
:::

::: {.column width="60%"}

@@

```{r}
knitr::include_graphics("assets/svm.png")
```
</div>
:::

@@

### We make use of pre-trained large language models
- Benefit from semantic similarity
- Text embeddings mean we can have an arbitrary number of classes
- Switch out classification target schema easily!

@@

- Here we do not say which skills are green or brown, we just use the `kmeans` algorithm to create two groups based on the embeddings:

![](assets/umap_1.png){fig-align="center" width=100%}

## Use clusters to classify new skills into the groups

# New terms to classify
new_terms = ["Sustainable textile recycling", "Coal burning technology"]

# Encode the new terms using the same model
new_embeddings = model.encode(new_terms)

# Transform the new embeddings using the already fit UMAP model
new_embeddings_umap = umap_model.transform(new_embeddings)

# Use the trained KMeans model to predict the cluster for the new embeddings
new_clusters = kmeans.predict(new_embeddings_umap)

# Mapping clusters to groups for readability
group_mapping = {1: 'Green Skills', 0: 'Brown Skills'}
classified_groups = [group_mapping[cluster] for cluster in new_clusters]

# Output the classification
for term, group in zip(new_terms, classified_groups):
    print(f"'{term}' is classified as: {group}")
```

- 'Sustainable textile recycling' is classified as: Green Skills

- 'Coal burning technology' is classified as: Brown Skills

## Visualize classification

![](assets/umap_2.png){fig-align="center" width=100%}

@@

Here the HISCO descriptions are in English, as are the translated titles from my biographies in Swedish.

@@

### 026 Metallurgists (HISCO description)

@@

### Karl Lund in 1923 (Input)

@@

## Keep titles in Swedish?

Instead of using the HISCO schema in English, I am using the SSYK classification from 2012.

Instead of using the multilingual OpenAI ADA model, I now use [The National Library of Sweden Swedish BERT model](https://huggingface.co/KB/bert-base-swedish-cased). It is trained on 200M sentences, 3000M tokens from books, news, government publications, swedish wikipedia and internet forums.

The input is from the Job Tech Data so we have a ground truth that they have labelled.

```{r}
knitr::include_url("https://huggingface.co/KB/bert-base-swedish-cased", height = 800)
```

## SSYK and Swedish BERT examples:

```{r}
library(tibble)
library(gt)

# New tibble with the provided data
new_occupation_data <- tibble(
    Rank = c(1, 2, 3, 4, 5),
    Closest_Match_Title_and_Code = c(
        "5326 - Ambulanssjukvårdare",
        "2226 - Ambulanssjuksköterskor m.fl.",
        "2227 - Geriatriksjuksköterskor",
        "5112 - Tågvärdar och ombordansvariga m.fl.",
        "2231 - Operationssjuksköterskor"
    )
)

# Create a gt table from the new tibble
new_occupation_table <- new_occupation_data %>%
    gt() %>%
    tab_header(
        title = md("**Closest SSYK Codes** to 'Ambulanssjukvårdare' (Code: 5326)")
    ) %>%
    cols_label(
        Rank = "Rank",
        Closest_Match_Title_and_Code = "Closest Match Title and Code"
    )

new_occupation_table %>% 
  tab_options(table.font.size = 40)

## SSYK and Swedish BERT examples:

```{r}
library(tibble)
library(gt)

# New tibble with the provided data
new_occupation_data <- tibble(
    Rank = c(1, 2, 3, 4, 5),
    Closest_Match_Title_and_Code = c(
        "3115 - Ingenjörer och tekniker inom kemi och kemiteknik",
        "2145 - Civilingenjörsyrken inom kemi och kemiteknik",
        "3113 - Ingenjörer och tekniker inom elektroteknik",
        "7321 - Prepresstekniker",
        "2163 - Planeringsarkitekter m.fl."
    )
)

# Create a gt table from the new tibble
new_occupation_table <- new_occupation_data %>%
    gt() %>%
    tab_header(
        title = md("**Closest Matches** to 'Processingenjör, kemiteknik (Code: 2512)'")
    ) %>%
    cols_label(
        Rank = "Rank",
        Closest_Match_Title_and_Code = "Closest Match Title and Code"
    )

new_occupation_table %>% 
  tab_options(table.font.size = 40)
```

@@
